---
title: "Analysis of Movie Success"
author: "Me"
date: "`r Sys.Date()`"
output: html_document
---

```{r message=FALSE}
library(regclass)
library(lubridate)

library(arules)
library(multcompView)
```

## Goals

We want to know what factors influence a movie's success.

We plan to use the following methods to look into this:

-   Looking at combinations of factors:

    -   Decision tree - this can handle both numeric and categorical data

        -   Numeric will make splits like: Budget > 10000
        -   Categorical will make splits like: Genre == "Horror"

    -   Market basket analysis - this can only handle categorical data

```{r}
movie <- read.csv("MyMoviesFall2022.csv", stringsAsFactors = TRUE)
dim(movie)
```

### Data Cleaning

Remove records with atypical/nonsensical values in:

-   `Budget`
-   `Theaters`

```{r}
hist(movie$Budget)
summary(movie$Budget)

movie <- movie[movie$Budget>0, ]

summary(movie$Theaters)

movie <- movie[movie$Theaters > 0, ]
```

Remove records with missing values in:

-   `Theaters`
-   `RunTime`

```{r}
# !=
summary(movie$RunTime)
summary(movie$Theaters)
summary(movie$InReleaseDays)
summary(movie$Success)
summary(movie$Rating)

movie <- movie[!is.na(movie$RunTime), ]
movie <- movie[!is.na(movie$Theaters), ]
movie <- movie[!is.na(movie$InReleaseDays), ]
movie <- movie[!is.na(movie$Success), ]

dim(movie)
```

Make `WeekOfMonth` categorical:

```{r}
movie$WeekOfMonth <- factor(movie$WeekOfMonth)
levels(movie$WeekOfMonth) <- c("First", "Second", "Third", "Fourth", "Fifth")

summary(movie$WeekOfMonth)
```

Make a column named `Month` that holds the name of the month from the `ReleaseDate` field.

```{r}
movie$Month <- month(mdy(movie$ReleaseDate), label = TRUE)
summary(movie$Month)
```

Drop the `ReleaseDate` field and `Movie` title columns:

```{r}
movie$ReleaseDate <- NULL
movie$Movie <- NULL
```

Investigate the following columns:

-   `Budget`
-   `RunTime`
-   `Theaters`

What's an issue with them? What are options to fix?

```{r}
# Making a copy to have a version that keeps some
# variables numeric
movie_copy <- movie
```

```{r}
# B - c(0, 1e6, 5e6, 1e7, 2e7, 3e7, 4e7, 5e7, 6e7, 7e7, 1e8, 5e8)
# R - c(60, 90, 95, 100, 105, 110, 120, 130, 150, 260)
# T - c(0, 700, 1500, 2000, 2500, 2800, 3100, 3500, 5000)

movie$Budget <- cut(
  movie$Budget,
  breaks = c(0, 1e6, 5e6, 1e7, 2e7, 3e7, 4e7, 5e7, 6e7, 7e7, 1e8, 5e8),
  include.lowest = TRUE
)
summary(movie$Budget)

movie$RunTime <- cut(
  movie$RunTime,
  breaks = c(60, 90, 95, 100, 105, 110, 120, 130, 150, 260),
  include.lowest = TRUE
)
summary(movie$RunTime)

movie$Theaters <- cut(
  movie$Theaters,
  breaks = c(0, 700, 1500, 2000, 2500, 2800, 3100, 3500, 5000),
  include.lowest = TRUE
)
summary(movie$Theaters)
```

Investigate the following columns:

-   `Rating`
-   `Distribution`
-   `TopBilledStar`
-   `MainGenre`
-   `SecondaryGenre`
-   `Director`

What's an issue with them? What are options to fix?

```{r}
# R - 20
# Dis - 75
# T - 25
# M - 80
# S - 80
# Dir - 10

movie$Rating <- combine_rare_levels(movie$Rating, threshold = 20)$values
movie$Distribution <- combine_rare_levels(movie$Distribution, threshold = 75)$values
movie$TopBilledStar <- combine_rare_levels(movie$TopBilledStar, threshold = 25)$values
movie$MainGenre <- combine_rare_levels(movie$MainGenre, threshold = 80)$values
movie$SecondaryGenre <- combine_rare_levels(movie$SecondaryGenre, threshold = 80)$values
movie$Director <- combine_rare_levels(movie$Director, threshold = 10)$values

# Rename new level to "Other"
levels(movie$Rating)[levels(movie$Rating) == "Combined"] <- "Other"

cols <- c(
  "Rating",
  "Distribution",
  "TopBilledStar",
  "MainGenre",
  "SecondaryGenre",
  "Director"
)

for (col in cols) {
  levels(movie[[col]])[levels(movie[[col]]) == "Combined"] <- "Other"
}
```

Drop hyphens from the following columns:

-   `Rating`
-   `Distribution`
-   `MainGenre`

Why? A package we're about to use hates hyphens.  Nothing theoretically fancy here; if we leave them in we get an error, take them out get results.

```{r}
# Example:
x <- c("a-b", "a-b-c", "-------z")
gsub("-", "", x, fixed = TRUE)
```

```{r}
movie$Rating <- gsub("-", "", movie$Rating, fixed = TRUE)
movie$Distribution <- gsub("-", "", movie$Distribution, fixed = TRUE)
movie$MainGenre <- gsub("-", "", movie$MainGenre, fixed = TRUE)
```

Change the `'-'` category in `Secondary` level to be `'None'`:

```{r}
# Example:
x <- factor(c("a", "-", "-", "-"))
summary(x)

levels(x)[levels(x) == "-"] <- "None"
summary(x)
```

```{r}
levels(movie$SecondaryGenre)[levels(movie$SecondaryGenre) == "-"] <- "None"
levels(movie$SecondaryGenre)
```

Convert the `Universe` column to say `'No'` if the `Universe` is listed as `'None'`; if not `'None'` say `'Yes'`:

```{r}
movie$Universe <- ifelse(movie$Universe == "None", "No", "Yes")
```

```{r}
movie$CriticRating <- NULL
```


Save the now cleaned data to a csv:

```{r}
dim(movie)
write.csv(movie, "mycleanmovies.csv", row.names = FALSE)
```

```{r}
movie$Budget <- movie_copy$Budget
movie$RunTime <- movie_copy$RunTime
movie$Theaters <- movie_copy$Theaters

write.csv(movie, "mycleanmoviesNUMs.csv", row.names = FALSE)
```

### Analysis

```{r}
movie <- read.csv("mycleanmovies.csv", stringsAsFactors = TRUE)
dim(movie)
```

Final prep

-   Remove columns used in definition of success
-   Remove columns that wouldn't be available at time of prediction
-   Remove redundant columns (budget \<-> est budget)

```{r}
keep_cols <- c(
  "Rating", "Distribution", "Budget",
  "RunTime", "TopBilledStar", "MainGenre",
  "SecondaryGenre", "WeekOfMonth", "Director",
  "Universe", "Theaters", "Success", "Month"
)

movie <- movie[, keep_cols]
dim(movie)
```

What percentage of movies are successful according to our definition of success?

```{r}
mean(movie$Success == 1) #DEFINING SUCCESS AS CRITIC RATING > 0.7
```

#### Single drivers

Lets start from the first column! How does success vary across ratings?

-   View with a `mosaic()`

```{r}
mosaic(Success ~ Rating, data = movie, equal = TRUE, inside = TRUE)
```

-   Create a summary table showing the rate of success by rating

    -   Rename the rate column to "ProbSuccess"

-   Create a summary table showing the number of records by rating

    -   Rename the count column to "Count"

-   Combine these two tables

```{r}
prob_success_by_rating <- aggregate(Success ~ Rating, data = movie, mean)
names(prob_success_by_rating)[2] <- "ProbSuccess"

n_by_rating <- aggregate(Success ~ Rating, data = movie, length)
names(n_by_rating)[2] <- "Count"

summary_table <- merge(prob_success_by_rating, n_by_rating, by = "Rating")
```

-   Perform an ANOVA test using `aov()`
-   Perform a follow up on the aov results using `TukeyHSD()`
-   Use `multcompLetters4()` to assign groups

```{r}
aov_results <- aov(Success ~ Rating, data = movie)
tukey_results <- TukeyHSD(aov_results)
tukey_letters <- multcompLetters4(aov_results, tukey_results)[["Rating"]]

letter_df <- data.frame(
  Rating = names(tukey_letters$Letters),
  Group = tukey_letters$Letters
)
```

-   Add the `letter_df` to our summary table

```{r}
summary_table <- merge(summary_table, letter_df, by = "Rating")
summary_table
```

-   Sort results from highest to lowest success

```{r}
summary_table <- summary_table[order(summary_table$ProbSuccess, decreasing = TRUE), ]
summary_table
```

Now repeat for every other column........

Very tedious! This is why programming exists!

Converting the code to a function:

```{r}
examine_movie_driver <- function(driver) {
  driver_formula <- formula(paste("Success ~", driver))

  # Find success rate per group
  success_rate <- aggregate(driver_formula, data = movie, FUN = function(x) mean(x==1))
  names(success_rate)[2] <- "ProbSuccess"

  # Find count per group
  record_count <- aggregate(driver_formula, data = movie, FUN = length)
  names(record_count)[2] <- "Count"

  # Combine success rates and counts
  success_summary <- merge(success_rate, record_count)

  # Perform anova and extract p value
  aov_results <- aov(driver_formula, data = movie)
  p_val <- anova(aov_results)$`Pr(>F)`[1]

  # Check if anova saw significant diff
  if (p_val < 0.05) {
    # Perform Tukey if significant diff
    tukey_results <- TukeyHSD(aov_results)
    tukey_letters <- multcompLetters4(aov_results, tukey_results)[[driver]]

    letter_df <- data.frame(
      names(tukey_letters$Letters),
      tukey_letters$Letters
    )
    names(letter_df) <- c(driver, "Group")

    # Combine tukey results to orignal summary
    success_summary <- merge(success_summary, letter_df)
  } else {
    # Give all same group if no sig diff
    success_summary$Group <- "a"
  }

  # Sort to have highest prob on top
  success_summary <- success_summary[order(success_summary$ProbSuccess, decreasing = TRUE), ]
  rownames(success_summary) <- NULL
  success_summary <- success_summary[, c(driver, "ProbSuccess", "Group", "Count")]

  # Create barplot (reordering factor levels to sort bars by rate)
  movie[, driver] <- factor(movie[, driver], levels = success_summary[, driver])
  mosaic(driver_formula, data = movie, inside = TRUE, equal = TRUE)

  success_summary
}
```

Using the function:

```{r}
examine_movie_driver("Rating")
```

```{r}
examine_movie_driver("Distribution")
```

Repeat for all possible drivers
  * Option 1 - copy / paste
  * Option 2 - loop (preferred route)

```{r}
for (n in names(movie)) {
  if (n != "Success") {
    print(examine_movie_driver(n))
  }
}
```

#### Combinations of drivers

##### Decision Tree

A decision tree is a nice way to explore drivers in a very interpretable way. If prediction accuracy is what you're after then decision trees often get outperformed by other methods (like a forest of decision trees!).

The parameters of the decision tree will play with are:

-   `cp` - "complexity parameter"

    -   based on the idea that a tree with a bunch of decisions is very complex
    -   the higher the cp, the fewer splits the tree will make
    -   the lower the cp, the more splits the tree will make
    -   its value means: the min improvement in the model needed to add a decision (so if the cp bar is too high no decisions added)

-   `minbucket`

    -   minimum number of observations in a 'leaf' node
    -   the higher the minbucket, the fewer splits the tree will make
    -   the lower the minbucket, the more splits the tree will make

Use `minbucket` of 50. Start with a low `cp` and adjust until you get about 5-8 splits (use `tree$cptable` to check this)

```{r}
tree <- rpart(Success ~ ., data = movie, cp = .01, minbucket = 50)

tree$cptable

summarize_tree(tree)
visualize_model(tree)
```

Note, the big focus on making categorical vars was for the market basket analysis. Decision trees can handle numeric data and it could actually make some of the decisions more readable in the viz.

```{r}
rpart_df <- read.csv("mycleanmoviesNUMs.csv", stringsAsFactors = TRUE)

rpart_cols <- c(
  "Rating", "EstimatedTotBudget", "RunTime", "MainGenre",
  "WeekOfMonth", "Month", "Success"
)

tree <- rpart(Success ~ ., data = rpart_df[, rpart_cols], cp = .007, minbucket = 100)
tree$cptable
tree
visualize_model(tree)
```

##### Market Basket Analysis

A somewhat creative approach to this problem is to think of all of these drivers as items in a shopping cart. What items are often 'bought' with Success?

If every time we see Tom Cruise in a cart we also see Success then Tom is a driver of success.

```{r}
# market basket analysis in R is verrry picky about data types
movie$Success <- ifelse(movie$Success == 1, "Success", "NotSuccess")
movie$Success <- factor(movie$Success)

movie_trans <- as(movie, "transactions")
# tomato tomato
mean(movie$Success == "Success")
itemFrequency(movie_trans)["Success=Success"]
```

```{r}
# All of arules is impossible to remember how to type up
# If you want to repeat this, keep examples of code around

# Parameters:
min_count <- 50 # Min number of movies to consider a rule
min_conf <- 0.25 # Min value of confidence
max_len <- 3 # How long can lhs of rule be (ie if "if a and b in basket then c" is len 2)

rules <- apriori(movie_trans,
  parameter = list(
    supp = min_count / length(movie_trans),
    conf = min_conf,
    maxlen = max_len
  ),
  appearance = list(default = "lhs", rhs = "Success=Success"),
  control = list(verbose = FALSE)
)

# {a & b -> c} is redundant with rule {b & a -> c}
rules <- rules[!is.redundant(rules)]
rules <- rules[is.significant(rules, movie_trans)]

rules_df <- DATAFRAME(rules)

rules_df <- rules_df[order(rules_df$confidence, decreasing = TRUE), ]
rules_df
```

```{r}
rules_cols <- c(
  "Rating", "Budget", "RunTime",
  "MainGenre", "WeekOfMonth", "Month", "Success"
)

movie_trans <- as(movie[, rules_cols], "transactions")

# Parameters:
min_count <- 20 # Min number of movies to consider a rule
min_conf <- 0.25 # Min value of confidence
max_len <- 3 # How long can lhs of rule be (ie if "if a and b in basket then c" is len 2)

rules <- apriori(movie_trans,
  parameter = list(
    supp = min_count / length(movie_trans),
    conf = min_conf,
    maxlen = max_len
  ),
  appearance = list(default = "lhs", rhs = "Success=Success"),
  control = list(verbose = FALSE)
)
rules <- rules[!is.redundant(rules)]
rules <- rules[is.significant(rules, movie_trans)]

rules_df <- DATAFRAME(rules)
rules_df <- rules_df[order(rules_df$confidence, decreasing = TRUE), ]
rules_df
```

Interpretation template:

-   Overall, the probability of success is `<itemFrequency(movie_trans)["Success=Success"]>`
-   Among movies with `<items in the lhs column>`, the probability of success increases by a factor of `<lift>` to `<confidence>`.
-   This is based on `<count>` movies.

Example:

```{r}
# itemFrequency(movie_trans)["Success=Success"]
# 0.1513542

# lhs                 rhs                support   confidence  coverage  lift      count
# {Budget=[0,1e+06]}  {Success=Success}  0.008497  0.888888    0.009559  5.872904  32
```

Overall, the probability of success is \~15%. Among movies with budgets \<= \$1M we see this probability of succcess increase by a factor of \~5.9 to be \~89%. This is based on 32 movies.

What does this mean in terms of how we defined success? Is this meaningful?
